---
title: "Project Checkpoint 1"
author: "Haylee Allen"
date: "4/7/2024"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Wages: Real Wage and Purchasing Power

The goal for this section is to Compare wages in retail to their real wage value. In doing so, we can reveal trends in wages and identify if wages are part of a larger issue or if wages have kept up with inflation enough to deduce that working in retail has paid the same over time. Data was provided via FRED: Wage data from https://fred.stlouisfed.org/series/CES4200000003 and CPI data from (https://fred.stlouisfed.org/series/CPIAUCSL

First, is data setup and cleanup. After loading the libraries, we look for null values and omit as necessary.

```{r data setup}
#load libraries
library(fredr)
library(tidyverse)
library(ggplot2)
library(dplyr)

#Get Data
projpath<-"C:/Users/dancu/Documents/GitHub/Software-Tools-Spring-2024/Data_Tables" #set directory
ddi <- read_ipums_ddi(file.path(projpath,"RealWage.csv"))
RealWageData<-read_ipums_micro(ddi)

#check for null values
sum(is.na(RealWageData))

#Data clean up
#create data to omit nulls
RealWageDataOmit1<-na.omit(RealWageData)
#Change the header names
colnames(RealWageDataOmit1)[2]<- "CPI"
colnames(RealWageDataOmit1)[3]<- "NominalWage"

#convert data to dates
RealWageDataOmit<-RealWageDataOmit1%>%
  mutate(DATE= as.Date(DATE, format = "%d/%m/%Y"))
```

Next, the real wage and purchasing power for each wage data point must be calculated. A new column is added for each. 

Real Wage in a year = (Nominal Wage in a Year/CPI in a Year)x100

```{r add new columns}
#Add in columns for purchasing power and real Wage
WageData<-RealWageDataOmit %>%
  mutate(
    RealWage = (NominalWage/CPI)*100)
```

Lastly,a plot of the data the data reveals trends which show  the real wage appears to have increased over time. 

```{r aplot}
#Plot real wage and purchasing power
ggplot(WageData, aes(x=DATE, y=RealWage, group = 1))+
  geom_line()+
  labs(title = "Real Wage", x = "Date", y = "Real Wage")
```
